\dispRity quick demo

### 2015-11-18 doc/dispRity-quick-demo.Rnw
demo update

This demo, which will use dummy data will quickly cover the following points:
\begin{enumerate}
\item 
\end{enumerate}

\section{Installation}
You can install this package easily if you are using the latest version of \R and \texttt{devtools}.

@@

\section{}

>>>>

This demo, which will use dummy data will quickly cover the following points:
\begin{enumerate}
\item Installing \dispRity
\item Ordinated matrices
\item Creating submatrices
\item Bootstrapping the data
\item Calculating disparity
\item Plotting the results
\item Testing differences
\end{enumerate}

\section{Installing \dispRity}
You can install this package easily if you are using the latest version of \R and \texttt{devtools}.

@@

\section{Ordinated matrices}
Ordinated matrices are transformed distance matrices based on various methods (e.g. MDS, PCA, PCO, PCoA). For this example, we will just generate \textit{dummy} data with 26 elements and 15 variables.

<<data>>=
## Loading the package
library(dispRity)
## random seed for reproducibility
set.seed(1)

## This is a multivariate matrix:
dummy_data <- replicate(15, rnorm(26, mean=sample(1:10)))
rownames(dummy_data) <- LETTERS
head(dummy_data[,1:5])
## We can the calculate the distance matrix:
dummy_dist <- dist(dummy_data)
## And ordinate it into 10 dimensions:
dummy_ord <- cmdscale(dummy_dist, k=10)
head(dummy_ord)
@ 
This ordinated data will constitute our multidimensional space (e.g. the ecospace, morphospace, cladistospace, etc...).

\section{Creating submatrices}
We can now create two different groups if we are interested in looking at the difference in occupancy of the multidimensional space between the vowels and the consonants.

<<cust.data>>=
## Creating a data.frame that contains information on the letter type
factors <- data.frame("type"=rep("consonant", 26), row.names=LETTERS,
    stringsAsFactors=FALSE)
factors[c(1,5,9,15,21,25),1] <- rep("vowels", 6)

## Splitting the data into sub-matrices according the the factors data.frame
data_series <- cust.series(dummy_ord, factors)
@

Note that from now on, objects are attributed the class \dispRity that summarises the object content.

<<class>>=
class(data_series)
str(data_series)
data_series
@

\section{Bootstrapping the data}
We can now bootstrap to get rid of problems from outliers and also rarefy the dataset to have the same number of vowels or consonants.

<<boot.data>>=
## Simple bootstrapping
boot_data <- boot.matrix(data_series)

## Bootstrapping and rarefaction to get the same number of elements
## in both categories (vowels and consonants)
rare_data <- boot.matrix(data_series, rarefaction=6)

## Again both objects can be quickly summarised by just calling them
boot_data
rare_data
@

\section{Calculating disparity}
We can calculate disparity as the sum of the variance of each dimension for each sub-matrix to measure the spread of the data for vowels of consonants.

<<disparity>>=
## Calculating disparity for both objects
disparity_boot <- dispRity(boot_data, metric=c(sum, variances))
disparity_rare <- dispRity(rare_data, metric=c(sum, variances))

## Again, printing the objects will just give basic information.
disparity_boot

## To actually get the results, we can use the summary function
summary(disparity_boot)
summary(disparity_rare)
@

\section{Plotting the results}
As we can guess from the summary, our results differ between consonants and vowels but consonants seems not affected by sample size.
But let's visualise it:

<<plot1, fig.width=10, fig.height=5, out.width='1\\linewidth'>>=
## Graphical options
quartz(width = 10, height = 5) ; par(mfrow = (c(1,2)), bty = "n")
## The disparity-through-time data in a time-binned way (with lines rather than
## boxes) with the observed data as crosses
plot(disparity_boot, main="raw data", elements=TRUE, ylab=c("disparity", "letters"))
## The disparity-through-time data in a time-sliced ways
plot(disparity_rare, main="rarefied data", elements=TRUE, ylab=c("disparity", "letters"))
@

\section{Testing differences}
Finally, to be sure if the disparity is different between vowels and consonants or not, we can apply a statistical test to both categories:

<<test>>=
## Graphical options
quartz(width = 10, height = 5) ; par(mfrow = (c(1,2)), bty = "n")
## The disparity-through-time data in a time-binned way (with lines rather than
## boxes) with the observed data as crosses
plot(disparity_boot, main="raw data", elements=TRUE, ylab=c("disparity", "letters"))
## The disparity-through-time data in a time-sliced ways
plot(disparity_rare, main="rarefied data", elements=TRUE, ylab=c("disparity", "letters"))
@
